\large How resource abundance and stochasticity affect organisms' range sizes

### 2024-07-26 writing/appendix-A-concepts-and-definitions.Rmd
small change

\noindent In viewing resources as a random variable, we not only obtain a formal framework for describing $\e(R)$, but also the spread around $\e(R)$, which we quantify with its variance. We use the notation $\var(R)$ to indicate the variance in $R$ around the mean, i.e., after accounting for changes in $\mu(t, u)$, and we use $\sigma^2(t, u)$ to indicate its function over time and time (with estimate $\widehat{\sigma^2}(t, u)$; see Fig. A\@ref(fig:multipanel-changing-var)). For example, the time at which berry bushes produce fruit may seem highly stochastic (i.e., unpredictable) to a young bear, but it becomes predictable as it learns to understand and predict how the mean amount of berry changes with the seasons. Still, whether next year will be a good or bad year for berries remains stochastic because it depends on factors the bear cannot predict (e.g., droughts, fires), but the bear can still expect $\sigma^2(t, u)$ to be lower in the winter since berries will be hard to find (i.e., $\mu(t, u) \approx 0$). Also note that, like with $\mu(t, u)$, what values $\sigma^2(t, u)$ can take depends on what distribution we assume $R$ follows. Figures A\@ref(fig:ex-gamma) and A\@ref(fig:ex-beta) show four examples of Gamma and Beta distributions, respectively, with different means and variances.

>>>>

\noindent In viewing resources as a random variable, we not only obtain a formal framework for describing $\e(R)$, but also the spread around $\e(R)$, which we quantify with its variance (i.e., stochasticity). We use the notation $\var(R)$ to indicate the variance in $R$ around the mean, i.e., after accounting for changes in $\mu(t, u)$, and we use $\sigma^2(t, u)$ to indicate its function over time and time (with estimate $\widehat{\sigma^2}(t, u)$; see Fig. A\@ref(fig:multipanel-changing-var)). For example, the time at which berry bushes produce fruit may seem highly stochastic (i.e., unpredictable) to a young bear, but it becomes predictable as it learns to understand and predict how the mean amount of berry changes with the seasons. Still, whether next year will be a good or bad year for berries remains stochastic because it depends on factors the bear cannot predict (e.g., droughts, fires), but the bear can still expect $\sigma^2(t, u)$ to be lower in the winter since berries will be hard to find (i.e., $\mu(t, u) \approx 0$). Also note that, like with $\mu(t, u)$, what values $\sigma^2(t, u)$ can take depends on what distribution we assume $R$ follows. Figures A\@ref(fig:ex-gamma) and A\@ref(fig:ex-beta) show four examples of Gamma and Beta distributions, respectively, with different means and variances.
